---
title: "Initial Korean covid 19"
author: "Maximiliano Ortiz"
date: "Fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
  
# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

When plotting the graphical results, I observed that the number of people alive are way higher than the number of people dead. Furthermore, when looking at the percentage chart, I found that in the chart of dead people 2/3 of It were men and 1/3 were women, suggesting that the disease mostly killed males more than females, while in the alive chart most of the alive people were women (about 57%) compared to men (about 43%). 

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results basically support the results we got in the graphical part. For Females, the percentage of women dying because of the disease is 1.4% while for men it is 3%, which is the double of the female percentage. This suggests that men have a higher chance of dying than women.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When observing the inferential results, I found out that the odds of men to die because of COVID in South Korea is 2.3 times greater than the women, which agrees to the past results, as we saw that most of the dead people because of COVID are mostly men. I also noticed that the P-value is low (0.0009), and thus this low value suggests that the chances of males and females dying because of COVID is not the same, because men have a higher chance of getting COVID and dying way more than women.

## Conclusions

The analysis of the South Korean COVID-19 patient data from January to May 2020 shows a clear association between sex and mortality due to COVID-19. Males had a significantly higher risk of dying from the disease compared to females. Specifically, men were about 2.3 times more likely to die than women, as supported by both graphical and numerical evidence. The chi-square and Fisher’s exact tests confirm this difference is statistically significant (p-value < 0.001), indicating that the chance of dying from COVID-19 is not independent of sex. These findings highlight important sex-related differences in COVID-19 outcomes that may be driven by biological, behavioral, or social factors. Further investigation into these factors could provide valuable insights to improve targeted interventions and healthcare strategies.

